---
title: "Midterm"
output: word_document
date: "2023-03-15"
---

# Packages
```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(nbastatR)
library(devtools)
library(cowplot)
```

# Creating Court 
```{r include=FALSE}
# Creating court and plotting
circle_points = function(center = c(0, 0), radius = 1, npoints = 360) {
  angles = seq(0, 2 * pi, length.out = npoints)
  return(data_frame(x = center[1] + radius * cos(angles),
                    y = center[2] + radius * sin(angles)))
}
# Court Dimenons & lines
width = 50
height = 94 / 2
key_height = 19
inner_key_width = 12
outer_key_width = 16
backboard_width = 6
backboard_offset = 4
neck_length = 0.5
hoop_radius = 0.75
hoop_center_y = backboard_offset + neck_length + hoop_radius
three_point_radius = 23.75
three_point_side_radius = 22
three_point_side_height = 14
# Court themes
court_themes = list(
  light = list(court = 'white',lines = 'black',text = '#f0ebd2'))
# Function to create court based on given dimensions
plot_court = function(court_theme = court_themes$light) {
  
  court_points = data_frame(
    x = c(width / 2, width / 2, -width / 2, -width / 2, width / 2),
    y = c(height, 0, 0, height, height),desc = "perimeter")
  
  court_points = bind_rows(court_points , data_frame(x = c(outer_key_width / 2, outer_key_width / 2, -outer_key_width / 2, -outer_key_width / 2),y = c(0, key_height, key_height, 0),desc = "outer_key"))
  
  court_points = bind_rows(court_points , data_frame(
    x = c(-backboard_width / 2, backboard_width / 2),
    y = c(backboard_offset, backboard_offset),
    desc = "backboard"))
  
  court_points = bind_rows(court_points , data_frame(
    x = c(0, 0), y = c(backboard_offset, backboard_offset + neck_length), desc = "neck"))
  
  foul_circle = circle_points(center = c(0, key_height), radius = inner_key_width / 2)
  
  foul_circle_top = filter(foul_circle, y > key_height) %>%
    mutate(desc = "foul_circle_top")
  
  foul_circle_bottom = filter(foul_circle, y < key_height) %>%
    mutate(
      angle = atan((y - key_height) / x) * 180 / pi,
      angle_group = floor((angle - 5.625) / 11.25),
      desc = paste0("foul_circle_bottom_", angle_group)
    ) %>%
    filter(angle_group %% 2 == 0) %>%
    select(x, y, desc)
  
  hoop = circle_points(center = c(0, hoop_center_y), radius = hoop_radius) %>%
    mutate(desc = "hoop")
  
  restricted = circle_points(center = c(0, hoop_center_y), radius = 4) %>%
    filter(y >= hoop_center_y) %>%
    mutate(desc = "restricted")
  
  three_point_circle = circle_points(center = c(0, hoop_center_y), radius = three_point_radius) %>%
    filter(y >= three_point_side_height, y >= hoop_center_y)
  
  three_point_line = data_frame(
    x = c(three_point_side_radius, three_point_side_radius, three_point_circle$x, -three_point_side_radius, -three_point_side_radius),
    y = c(0, three_point_side_height, three_point_circle$y, three_point_side_height, 0),
    desc = "three_point_line"
  )
  
  court_points = bind_rows(
    court_points,
    foul_circle_top,
    foul_circle_bottom,
    hoop,
    restricted,
    three_point_line
  )
  
  
  court_points <- court_points
  
  # Final plot creation
  ggplot() + geom_path(data = court_points,aes(x = x, y = y, group = desc),
  color = court_theme$lines) +
  coord_fixed(ylim = c(0, 45), xlim = c(-25, 25)) + 
  theme_minimal(base_size = 22) + 
  theme(text = element_text(color = court_theme$text),     
      plot.background = element_rect(fill = '#00471b', color = '#00471b'),   panel.background = element_rect(fill = court_theme$court, color = court_theme$court),   panel.grid = element_blank(),
      panel.border = element_blank(),                                                          axis.text = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      legend.background = element_rect(fill = court_theme$court, color = court_theme$court),
      legend.margin = margin(0, 0, 0, 0, unit = "lines"),
      legend.position = "bottom",
      legend.key = element_blank(),
      legend.text = element_text(size = rel(1.0))
    )
}
```

#2014
```{r echo=FALSE}
Sys.setenv(VROOM_CONNECTION_SIZE=400000)
# Grab team names to type into teams_shots function
nba_teams() %>% filter(isNonNBATeam == 0)
# Grab team shot data
bucks14 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2014, season_types = "Regular Season")
# Clean data to fit court dimensions
bucks14 <- bucks14 %>% mutate(x = as.numeric(as.character(locationX)) / 10, y = as.numeric(as.character(locationY)) / 10 + hoop_center_y)
#Horizontally Flip the Data
bucks14$x <- bucks14$x * -1 
#Creating the Chart
p1 <- plot_court(court_themes$light) + 
  geom_point(data = bucks14, aes(x = x, y = y, color = isShotMade, fill = isShotMade),alpha = .5, size = 1.5, shape = 21, stroke = .5) +  
  scale_color_manual(values = c("#009429","#F7A500"), aesthetics = "color", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_fill_manual(values = c("#009429","#F7A500"), aesthetics = "fill", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_x_continuous(limits = c(-27.5, 27.5)) +
  scale_y_continuous(limits = c(0, 40)) +
  theme(plot.title = element_text(hjust = .5, size = 22, face = "bold", vjust = 0),
        plot.subtitle = element_text(hjust = .5, size = 10, face = "bold", vjust = 1),
        legend.position = c(.475, .9),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(hjust = .5, size = 10, face = "bold", colour = "black")) +
  ggtitle(label = "Milwaukee Bucks", subtitle = "2013-14 Regular Season (W-L: 15-67 W/L%: .183)")
ggdraw(p1) + theme(plot.background = element_rect(fill="#00471b", color = NA)) 
ggsave("bucks14.png", height = 6, width = 6, dpi = 300)
```
•	Worst record in NBA
•	3PA/FGA: 1552/6737 (.230)
•	FG%: .438 (26th)	
•	3P%: .353 (20th)	
•	2P%: .464 (28th)

#2017
```{r echo=FALSE}
Sys.setenv(VROOM_CONNECTION_SIZE=400000)
# Grab team names to type into teams_shots function
nba_teams() %>% filter(isNonNBATeam == 0)
# Grab team shot data
bucks17 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2017, season_types = "Regular Season")
# Clean data to fit court dimensions
bucks17 <- bucks17 %>% mutate(x = as.numeric(as.character(locationX)) / 10, y = as.numeric(as.character(locationY)) / 10 + hoop_center_y)
# Horizontally flip the data
bucks17$x <- bucks17$x * -1 
#Creating the Chart
p1 <- plot_court(court_themes$light) +
  geom_point(data = bucks17, aes(x = x, y = y, color = isShotMade, fill = isShotMade),alpha = .5, size = 1.5, shape = 21, stroke = .5) +  
  scale_color_manual(values = c("#009429","#F7A500"), aesthetics = "color", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_fill_manual(values = c("#009429","#F7A500"), aesthetics = "fill", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_x_continuous(limits = c(-27.5, 27.5)) +
  scale_y_continuous(limits = c(0, 40)) +
  theme(plot.title = element_text(hjust = .5, size = 22, face = "bold", vjust = 0),
        plot.subtitle = element_text(hjust = .5, size = 10, face = "bold", vjust = 1),
        legend.position = c(.475, .9),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(hjust = .5, size = 10, face = "bold", colour = "black")) +
  ggtitle(label = "Milwaukee Bucks", subtitle = "2016-17 Regular Season (W-L: 42-40 W/L%: .512)")
ggdraw(p1) + theme(plot.background = element_rect(fill="#00471b", color = NA)) 
ggsave("bucks17.png", height = 6, width = 6, dpi = 300)
```
•	Two games above .500
•	3PA/FGA: 1946/6715 (.290)
•	FG%: .474 (4th)	
•	3P%: .370 (10th)	
•	2P%: .516 (6th)

#2019
```{r echo=FALSE}
Sys.setenv(VROOM_CONNECTION_SIZE=400000)
# Grab team names to type into teams_shots function
nba_teams() %>% filter(isNonNBATeam == 0)
# Grab team shot data
bucks19 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2019, season_types = "Regular Season")
# Clean data to fit court dimensions
bucks19 <- bucks19 %>% mutate(x = as.numeric(as.character(locationX)) / 10, y = as.numeric(as.character(locationY)) / 10 + hoop_center_y)
# Horizontally flip the data
bucks19$x <- bucks19$x * -1 
#Creating the Chart
p1 <- plot_court(court_themes$light) +
  geom_point(data = bucks19, aes(x = x, y = y, color = isShotMade, fill = isShotMade),alpha = .5, size = 1.5, shape = 21, stroke = .5) +  
  scale_color_manual(values = c("#009429","#F7A500"), aesthetics = "color", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_fill_manual(values = c("#009429","#F7A500"), aesthetics = "fill", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_x_continuous(limits = c(-27.5, 27.5)) +
  scale_y_continuous(limits = c(0, 40)) +
  theme(plot.title = element_text(hjust = .5, size = 22, face = "bold", vjust = 0),
        plot.subtitle = element_text(hjust = .5, size = 10, face = "bold", vjust = 1),
        legend.position = c(.475, .9),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(hjust = .5, size = 10, face = "bold", colour = "black")) +
  ggtitle(label = "Milwaukee Bucks", subtitle = "2018-19 Regular Season (W-L: 60-22 W/L%: .732)")
ggdraw(p1) + theme(plot.background = element_rect(fill="#00471b", color = NA)) 
ggsave("bucks19.png", height = 6, width = 6, dpi = 300)
```
•	Best record in NBA
•	3PA/FGA: 3134/7471 (.419)
•	FG%: .476 (3rd)	
•	3P%: .353 (15th)	
•	2P%: .565 (1st)

# 2021
```{r echo=FALSE}
Sys.setenv(VROOM_CONNECTION_SIZE=400000)
# Grab team names to type into teams_shots function
nba_teams() %>% filter(isNonNBATeam == 0)
# Grab team shot data
bucks21 <- teams_shots(teams = "Milwaukee Bucks", seasons = 2021, season_types = "Regular Season")
# Clean data to fit court dimensions
bucks21 <- bucks21 %>% mutate(x = as.numeric(as.character(locationX)) / 10, y = as.numeric(as.character(locationY)) / 10 + hoop_center_y)
# Horizontally flip the data
bucks21$x <- bucks21$x * -1 
#Creating the chart
p1 <- plot_court(court_themes$light) +
  geom_point(data = bucks21, aes(x = x, y = y, color = isShotMade, fill = isShotMade),alpha = 0.5, size = 1.5, shape = 21, stroke = .5) +  
  scale_color_manual(values = c("#009429","#F7A500"), aesthetics = "color", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_fill_manual(values = c("#009429","#F7A500"), aesthetics = "fill", breaks=c("TRUE", "FALSE"), labels=c("Made", "Missed")) +
  scale_x_continuous(limits = c(-27.5, 27.5)) +
  scale_y_continuous(limits = c(0, 40)) +
  theme(plot.title = element_text(hjust = .5, size = 22, face = "bold", vjust = 0),
        plot.subtitle = element_text(hjust = .5, size = 10, face = "bold", vjust = 1),
        legend.position = c(.475, .9),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(hjust = .5, size = 10, face = "bold", colour = "black")) +
  ggtitle(label = "Milwaukee Bucks", subtitle = "2020-21 Regular Season (W-L: 46-26 W/L%: .639)")
ggdraw(p1) + theme(plot.background = element_rect(fill="#00471b", color = NA)) 
ggsave("bucks21.png", height = 6, width = 6, dpi = 300)
```
•	Won NBA Championship
•	Regular season was shortened to 72 games
•	3PA/FGA: 2669/6610 (.404)
•	FG%: .487 (3rd)	
•	3P%: .389 (5th)	
•	2P%: .554 (4th)






2013-14
The 2013-14 Milwaukee Bucks won only 15 games for the worst record in the NBA. The shot chart shows that this team shot more mid-range shots than the other seasons. The team ranked 24th in 3PA, and 11th in 2PA. Only 23% of their shots were three-pointers. Compared to the rest of the league, they were better three-point shooters than two-point shooters. They had the third worst two-point percentage in the league, probably because of the choice to shoot long twos, rather than cutting to the basket or stepping back for a three. 

2016-17
This year the Bucks continued being a mid-table team by finishing with a record of 42-40. Between the 2014-15 season and the 2017-18 season, the Bucks finished with 41, 33, 42, and 44 wins, losing in the first round of the playoffs three times. They shot about 400 more threes than in 2013-14, shooting a three-pointer 29% of the time. Taking better shots increased their shooting percentages as they ranked in the top ten in FG, 3P, and 2P. 

2018-19
The Bucks finished the year with the best record in the NBA of 60-22 just five years after their worst season in franchise history. This was the first year for Mike Budenholzer as head coach and it is clear from the shot chart that there was a shift. They shot twice as many threes this season than in 2013-14, shooting a three 41.9% of the time. The team was completely average when it came to three-point percentage, but there two-point percentage was the best in the league with the combination of Giannis Antetokounmpo’s MVP season of scoring in the paint and Khris Middleton’s mid-range scoring.

2020-21
This year’s regular season was shortened by 10 games due to Covid-19. Milwaukee finished with a 46-26 record and went on to beat the Phoenix Suns in the Finals to win the NBA Championship. Three-point attempts were similar to recent years at around 40% of shots. The team ranked in the top five for FG, 3P, and 2P percentages. The offense was very well-rounded as they ranked in the top half of the league for almost all offensive categories.

